---
title: "Cafe Bazaar Report"
author: "Mohammad Mazraeh"
date: "May 14, 2016"
output: html_document
---
# Synposis
In this research we want to analyse user payement data and cluster users in order to plan for promotions or solutions to make more money (Increase users long term value)!
```{r,echo=FALSE}
suppressMessages(library(dplyr, warn.conflicts = FALSE, quietly=TRUE))
suppressMessages(library(ggplot2, warn.conflicts = FALSE, quietly=TRUE))
suppressMessages(library(grid, warn.conflicts = FALSE, quietly=TRUE))
suppressMessages(library("data.table",warn.conflicts = FALSE, quietly=TRUE))
# library(ggplot2)
library(grid)
# library(dplyr)
# library(data.table)
# library(grid)
```
# Data Processing
### Generating Data
In this part we would simulate some data as input for our analysis.  
<p>
<li>_**n**_ is number of data to generate.</li>
<li>_**numUsers**_ is the number of users.</li>
<li>_**typeFactor**_ is a factor including software types. in this analysis it's App or Game.</li>
</p>
We assume the data table would have one ID and 3 features:  
_**type**_:  the payement was for an Application or a Game</br>
_**amount**_:  the amount of payement</br>
_**OrderDate**_:  payement date</br>
It is assumed that installations with no payement are recorderd as _amount_ = 0.
Also any payement is recorderd in the table.
```{r}
n <- 3000
numUsers <- 500
typeFactor <- as.factor(c('App','Game'))
```
Let's generate the data:</br>
for each record _CustID_ is sampled from user IDs. _type_ is set to App or Game with eual probability.for _amount_ variable 0.9 of records will have zero value (Installations with no payement) and others are assumed to be from a normal distribution with arbitary mean and std. at last _orderDate_ is sampled randomly between "2013/01/01" and "2016/01/01".
```{r}
orders <- 
      data.table(CustID = sample(1:numUsers,size = n,replace = TRUE),
                 type = sample(typeFactor,n,replace = TRUE, prob = c(0.5,0.5)),
                 amount = round(sample(c(abs(rnorm(n = 0.1*n,mean = 3000,sd = 2000)),rep(0,0.9*n))),-2),
                 OrderDate = sample(seq(
                       as.Date('2013/01/01'), 
                       as.Date('2016/01/01'), 
                       by="day"), n,replace = TRUE))
```
### Convert Data to RFM Model
One of the most useful techniques to analyse user purchase history is to use RFM Models. R,F and M stand for Recency,Frequency and Monetary respectively. In this part we want to convert our input data into RFM Features:  
<li>_**FirstPurchaseDate**_: </li>
<li>_**LastPurchaseDate**_: </li>
<li>_**NumberofOrders**_: </li>
<li>_**TypeCount**_: </li>
<li>_**TotalAmount**_: </li>
```{r}
# Building RFM Features
RFM <- orders  %>% 
      group_by(CustID)  %>% 
      summarise(FirstPurchaseDate = min(OrderDate),
                LastPurchaseDate = max(OrderDate),
                NumberOrders = n(),
                TypeCount = length(unique(type)) ,
                TotalAmount = sum(amount))

LastDate <- max(RFM$LastPurchaseDate)
RFM <- RFM %>% mutate(R = as.numeric(LastDate-LastPurchaseDate))

```
#Analysis
##  Exploratory Data Analysis 
```{r,echo=FALSE,warning=FALSE,error=FALSE}
PlotRecency <- ggplot(RFM, aes(R %/% 7)) + geom_histogram(bins = 30) 
PlotFreq <- ggplot(RFM, aes(NumberOrders)) + geom_histogram()
PlotMonetary <- ggplot(RFM, aes(TotalAmount)) + geom_histogram()
PlotBreadth <- ggplot(RFM, aes(TypeCount)) + geom_histogram()
PlotTenure <- ggplot(RFM, aes(as.numeric(LastDate - FirstPurchaseDate) %/% 7)) + geom_histogram()
grid.newpage()
pushViewport(viewport(layout = grid.layout(3, 2)))
print(PlotRecency, vp = viewport(layout.pos.row = 1,layout.pos.col = 1))
print(PlotFreq, vp = viewport(layout.pos.row = 1,layout.pos.col = 2))
print(PlotMonetary, vp = viewport(layout.pos.row = 2,layout.pos.col = 1))
print(PlotBreadth, vp = viewport(layout.pos.row = 2,layout.pos.col = 2))
print(PlotTenure, vp = viewport(layout.pos.row = 3,layout.pos.col = 1))
```

```{r, echo=FALSE}
RFM_Segs <- data.table(Recency_Week = as.numeric(LastDate - RFM$LastPurchaseDate) %/% 7)
RFM_Segs$Recency <- ordered(ifelse(RFM_Segs$Recency_Week <= 30,"1-30",
                                   ifelse(RFM_Segs$Recency_Week <= 50,"30-50",
                                          ifelse(RFM_Segs$Recency_Week <= 60,"50-60",
                                                 ifelse(RFM_Segs$Recency_Week <= 90,"60-90","90+")))),
                            levels = c("1-30","30-50","50-60","60-90","90+"))

RFM_Segs$Frequency_count <- RFM$NumberOrders
RFM_Segs$Frequency <- ordered(ifelse(RFM_Segs$Frequency <= 1,"0-1",
                                     ifelse(RFM_Segs$Frequency <= 4,"2-4",
                                            ifelse(RFM_Segs$Frequency <= 8,"5-8",
                                                   ifelse(RFM_Segs$Frequency <= 10,"9-10","10+")))),
                              levels = c("0-1","2-4","5-8","9-10","10+"))
RFM_Segs$Monetary_Value <- RFM$TotalAmount
RFM_Segs$Monetary <- ordered(ifelse(RFM_Segs$Monetary_Value <= 0,"0",
                                     ifelse(RFM_Segs$Monetary_Value <= 1500,"0-1.5K",
                                            ifelse(RFM_Segs$Monetary_Value <= 5000,"1.5K-5K",
                                                   ifelse(RFM_Segs$Monetary_Value <= 10000,"5K-10K","10K+")))),
                              levels = c("0","0-1.5K","1.5K-5K","5K-10K","10K+"))


RFM_Segs$Breadth_count <- RFM$TypeCount
RFM_Segs$Breadth <-  ordered(ifelse(RFM_Segs$Breadth_count <= 1,"1","2"),
                             levels = c("1","2"))

RFM_Segs$Tenure_weeks <- as.numeric(LastDate - RFM$FirstPurchaseDate) %/% 7
RFM_Segs$Tenure <- ordered(ifelse(RFM_Segs$Monetary_Value <= 2,"0-2",
                                  ifelse(RFM_Segs$Monetary_Value <= 10,"2-10",
                                         ifelse(RFM_Segs$Monetary_Value <= 75,"10-75",
                                                ifelse(RFM_Segs$Monetary_Value <= 110,"75-110","110+")))),
                           levels = c("0-2","2-10","10-75","75-110","110+"))

```
```{r,echo=FALSE}
P1 <- ggplot(RFM_Segs, aes(Recency,Frequency))
P1 <- P1 +  geom_point(aes(size = Monetary_Value,colour = 'pink'))
print(P1)
```